---
title: "basic_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
 
concept of tidy data
 
Rstudio.cloud
 
create a project folder
create a data folder
upload a csv 
load in libraries
```{r}
library(tidyverse)


```


check the working directory, make sure it is set to the right place

```{r}
getwd()
s
```


load csv into a dataframe
```{r}
setwd("/cloud/project/basic_project_template/data")

df <- read_csv('SAFI_clean.csv')

```
```{r}

spec(df)
```


explore the dataframe
```{r}

head(df)


```

```{r}

names(df)


```


change the column names

```{r}

names(df)[7]


```

```{r}

names(df)[7] <- "number_of_rooms"


```

check to makesure it worked 

```{r}

names(df)


```

change a column type - 
            factors to numeric
            factors to string
```{r}

names(df)


```

use the pipe %>% and select a group of columns
```{r}

names(df)


```

use the pipe %>% filter a column
```{r}

names(df)


```

use unique to understand the varaiables in a column 
```{r}

names(df)


```


create a new column using mutate
```{r}

names(df)


```
create a new column using mutate and if else
```{r}

names(df)


```
group by and summarize ( the pivot table)
```{r}

names(df)


```
create a visualization 
  histogram
  scatter plot

```{r}

names(df)


```
how to join a table

```{r}

names(df)


```
put it together with R markdown 
  render a document
```{r}

names(df)


```

            